Theory # Portfolio

```{r}
library(shinydashboard)
library(shinyWidgets)
library(broom)

colnamer <- function(data, name) {
    data %>%
        `colnames<-`(name)
}
```

- The Sharpe Ratio is a measure of the return vs the risk ratio of a portfolio. 
- The Capital Asset Pricing Model (CAPM) will be used to calculate market beta for our assets and portfolio 
- The Fama-French Multi-Factor model will be used to explore multiple linear regression

For more, we can look at 

- (Asset Prices: A Theory of Market Equilibrium under Conditions of Risk)[https://www.efalken.com/LowVolClassics/sharpe64.pdf]
- (The Sharpe Ratio)[https://web.stanford.edu/~wfsharpe/art/sr/sr.htm]
- (Common Risk Factors in the Returns on Stocks and Bonds)[https://rady.ucsd.edu/faculty/directory/valkanov/pub/classes/mfe/docs/fama_french_jfe_1993.pdf]

# Sharpe Ratio

The Sharpe Ratio is defined as the mean of the excess monthly portfolio returns above the risk-free rate, divided by the standard deviation of the excess monthly portfolio returns above the risk-free rate. This is the formulation for 1994 as opposed to The. 

1966 SHARPE Ratio measures excess returns per unit of risk, where we take the standard deviation to represent portfolio risk. 

Sharpe Ratio = $\bar{R_p - R_f} / \sigma_{excess}$

## Sharpe Ratio in xts

```{r}
# choose a risk free rate
rfr <- .0003

# get sharpe ratio
(sharpe_xts <- SharpeRatio(portfolio_returns_xts_rebalanced_monthly,
                          Rf = rfr,
                          FUN = "StdDev") %>%
    `colnames<-`("sharpe_xts"))
```

# Sharpe Ratio in the Tidyverse

```{r}
sharpe_tidy <- portfolio_returns_dplyr %>%
    summarise(sharpe = mean(returns - rfr) / sd(returns - rfr))
```

# Sharpe Ratio in Tidyquant

```{r}
sharpe_tq <- portfolio_tq %>%
    tq_performance(Ra = returns,
                   performance_fun = SharpeRatio,
                   Rf = rfr,
                   FUN = "StdDev") %>%
    colnamer("sharpe_tq")
```

Now we can compare our Sharpe ratio to that of the S&P 500 in the same time period 

```{r}
getSymbols("SPY",
           src = "yahoo",
           from = "2012-12-31",
           auto.assign = TRUE,
           warnings = FALSE) %>%
    map(., ~ Ad(get(.x))) %>%
    reduce(merge) %>%
    colnamer("SPY") %>%
    to.monthly(indexAt = "lastof",
               OHLC = FALSE) -> market_returns_xts

market_returns_xts %>%
    tk_tbl(preserve_index = TRUE,
           rename_index = "date") %>%
    mutate(returns = (log(SPY) - log(lag(SPY)))) %>%
    na.omit() %>%
    summarise(ratio = mean(returns - rfr) / sd(returns - rfr)) -> market_sharpe
```

Our portfolio has underperformed the market during the chosen period.

# Visualizing Sharpe Ratio

Before we visualize the actual sharpe ratio, we can get a sense for what proportion of our portfolio returns exceeded the RFR.

```{r}
portfolio_tq %>%
    mutate(ratio = mean(returns - rfr) / sd(returns - rfr),
           returns_below_rfr = if_else(returns < rfr,
                                       returns,
                                       as.numeric(NA)),
           returns_above_rfr = if_else(returns > rfr,
                                       returns,
                                       as.numeric(NA))) %>%
    mutate_if(is.numeric, ~ round(.x, 4)) -> sharpe_returns
```

```{r}
sharpe_returns %>%
    ggplot(aes(x = date)) +
    geom_point(aes(y = returns_below_rfr),
               color = "orange") +
    geom_point(aes(y = returns_above_rfr),
               color = "forestgreen") +
    geom_vline(xintercept = as.numeric(as.Date("2016-11-30")),
               color = "blue",
               lty = 2,
               alpha = 0.2) +
    geom_hline(yintercept = rfr,
               color = "mediumpurple",
               lty = 5,
               alpha = 0.2) +
    annotate(geom = "text",
             x = as.Date("2016-11-30"),
             y = - 0.04,
             label = "Election",
             fontface = "plain",
             angle = 90,
             alpha = 0.5,
             vjust = 1.5) +
    ylab("% Monthly Returns")
```

Next we can build a histogram of the distribution of returns 

```{r}
sharpe_returns %>%
    ggplot(aes(x = returns)) +
    geom_histogram(alpha = 0.45,
                   binwidth = 0.01,
                   fill = "cornflowerblue",
                   color = "black") +
    geom_vline(xintercept = rfr,
               color = "mediumpurple",
               lty = 2,
               alpha = 0.3) +
    annotate(geom = "text",
             x = rfr,
             y = 13,
             label = "rfr",
             fontface = "plain",
             angle = 90,
             alpha = 0.5,
             vjust = 1)
```

We can also view our portfolio's Sharpe Ratio in comparison with other assets.

```{r}
asset_returns_long %>%
    summarize(std_dev = sd(returns),
              sharpe = mean(returns - rfr) / sd(returns - rfr)) %>%
    add_row(asset = "Portfolio",
            std_dev = port_rolling_sd_xts[1],
            sharpe = sharpe_tq$sharpe_tq) %>%
    ggplot(aes(x = std_dev,
               y = sharpe,
               color = asset)) +
    geom_point(size = 2) +
    geom_text(aes(x = sd(portfolio_tq$returns),
                  y = sharpe_tq$sharpe_tq + 0.02,
                  label = "Portfolio")) +
    ylab("Sharpe Ratio") +
    xlab("Standard Deviation") +
    ggtitle("Sharpe Ratio vs Standard Deviation")
```

S&P still beats our portfolio, but it seems to have more _very slightly_ more risk.

These overall numbers might obscure important periods of fluctuation in our data.

# Rolling Sharpe Ratio in xts

```{r}
window <- 24

rollapply(portfolio_returns_xts_rebalanced_monthly,
          window,
          function(x) SharpeRatio(x,
                           Rf = rfr,
                           FUN = "StdDev")) %>%
    na.omit() %>%
    colnamer("xts") -> roll_sharpe_xts
```

# Rolling Sharpe Ratio with the Tidyverse and tibbletime

```{r}
# create rolling function
sharpe_roll_24 <- rollify(
    function(returns) {
        ratio <- mean(returns - rfr) / sd(returns - rfr)
    },
    window = window)

portfolio_returns_dplyr %>%
    as_tbl_time(index = date) %>%
    mutate(sharpe = sharpe_roll_24(returns)) %>%
    na.omit() %>%
    select(-returns) -> roll_sharpe_tidy
```

# Rolling Sharpe Ratio with Tidyquant

```{r}
sharpe_tq_roll <- function(df) {
    SharpeRatio(df,
                Rf = rfr,
                FUN = "StdDev")
}

portfolio_tq %>%
    tq_mutate(select = returns,
              mutate_fun = rollapply,
              width = window,
              align = "right",
              FUN = sharpe_tq_roll,
              col_rename = "tq_sharpe") %>%
    na.omit() -> roll_sharpe_tq
```

# Visualizing the Rolling Sharpe Ratio

```{r}
highchart(type = "stock") %>%
    hc_title(text = "Rolling 24 Month Sharpe") %>%
    hc_add_series(roll_sharpe_xts,
                  name = "Sharpe",
                  color = "mediumpurple") %>%
    hc_navigator(enabled = FALSE) %>%
    hc_scrollbar(enabled = FALSE) %>%
    hc_add_theme(hc_theme_flat()) %>%
    hc_exporting(enabled = TRUE)
```

In ggplot

```{r}
roll_sharpe_xts %>%
    tk_tbl(preserve_index = TRUE,
           rename_index = "date") %>%
    rename(rolling_sharpe = xts) %>%
    ggplot(aes(x = date,
               y = rolling_sharpe)) +
    geom_line(color = "mediumpurple") +
    ggtitle("Rolling 24 Month Sharpe Ratio") +
    labs(y = "Rolling Sharpe Ratio")
```

# Shiny App Sharpe Ratio

```{r}
symbols <- c("SPY", "EFA", "IJS", "EEM", "AGG")

make_input <- function(symbol, index) {
    fluidRow(
        column(6,
               textInput(paste0("stock_", symbol),
                         paste("Stock", index),
                         symbol)),
        column(6,
               numericInput(paste0("wt_", index),
                            "Portfolio %",
                            value = 20,
                            min = 0,
                            max = 100)))
}

stock_inputs <- imap(symbols, ~ make_input(.x, .y))

date_input <- fluidRow(
    column(6,
           dateInput("date",
                     "Starting Date",
                     "2013-01-01",
                     format = "yyyy-mm-dd")),
    column(6,
           numericInput("window",
                        "Window",
                        min = 1,
                        max = 24,
                        value = 6)),
        column(6,
               numericInput(
                   "rfr",
                   "RFR %",
                   min = .01,
                   max = 1,
                   value = 0.03,
                   step = 0.01)))

buttons <- actionButton("go",
                        "Submit",
                        class = "btn btn-success")

ui <- fluidPage(
    theme = shinythemes::shinytheme("simplex"),
    sidebarLayout(
    sidebarPanel(!!!stock_inputs,
                 date_input,
                 buttons,
                 width = 2),
    mainPanel(fluidRow(highchartOutput("portfolio_sharpe",
                                       height = "500px",
                                       width = "900px")),
              fluidRow(
                  useShinydashboard(),
                  valueBoxOutput("box_portfolio"),
                  valueBoxOutput("box_market")))
))
```

```{r}
get_returns <- function(sym, sym_name, date) {
            getSymbols(sym,
                   src = "yahoo",
                   from = date,
                   to = "2018-01-01",
                   auto.assign = TRUE,
                   warnings = FALSE) %>%
            map(., ~ Ad(get(.x))) %>%
            reduce(merge) %>%
            colnamer(sym_name) %>%
            to.monthly(indexAt = "lastof",
                       OHLC = FALSE) %>%
            Return.calculate(method = "log") %>%
            na.omit()
}

server <- function(input, output, session) {
    # calculate market returns based on starting date
    market_returns <- eventReactive(input$go, {
    get_returns("SPY", "SPY", input$date)})

    # calculate overall and rolling market Sharpe ratio
    market_sharpe <- eventReactive(input$go, {
        SharpeRatio(market_returns(),
                    Rf = input$rfr,
                    FUN = "StdDev")})

    market_rolling_sharpe <- eventReactive(input$go, {
        rollapply(market_returns(),
                  input$window,
                  function(x) SharpeRatio(x,
                                   Rf = input$rfr,
                                   FUN = "StdDev")) %>%
            na.omit()})

    # get portfolio returns
    portfolio_returns <- eventReactive(input$go, {
        # grab inputs
        syms <- c(input$stock_SPY, input$stock_EFA,
                  input$stock_IJS, input$stock_EEM,
                  input$stock_AGG)

        # validate weights
        validate(need(input$wt_1 + input$wt_2 +
                     input$wt_3 + input$wt_4 +
                     input$wt_5 == 100,
                     "Portfolio weights must sum to 100%!"))
        
        weights <- c(input$wt_1 / 100, input$wt_2 / 100,
                     input$wt_3 / 100, input$wt_4 / 100,
                     input$wt_5 / 100)

        get_returns(syms, syms, input$date) %>%
            Return.portfolio(weights = weights)})

    # calculate portfolio Sharpe ratio
    portfolio_sharpe <- eventReactive(input$go, {
        # validate weights
        validate(need(input$wt_1 + input$wt_2 +
                      input$wt_3 + input$wt_4 +
                     input$wt_5 == 100,
                     "Portfolio weights must sum to 100%!"))

        SharpeRatio(portfolio_returns(),
                    Rf = input$rfr,
                    FUN = "StdDev")})

    # calculate portfolio rolling Sharpe ratio
    portfolio_rolling_sharpe <- eventReactive(input$go, {
        rollapply(portfolio_returns(),
                  input$window,
                  function(x) SharpeRatio(x,
                                   Rf = input$rfr,
                                   FUN = "StdDev")) %>%
            na.omit()})

    # plot 
    output$portfolio_sharpe <- renderHighchart({
        validate(need(input$go,
                      "Please choose your portfolio assets, weights, rfr, rolling window and start date and click submit."))

        highchart(type = "stock") %>%
            hc_title(text = "Rolling Sharpe") %>%
            hc_add_series(portfolio_rolling_sharpe(),
                          name = "Portfolio",
                          color = "mediumpurple") %>%
            hc_add_series(market_rolling_sharpe(),
                          name = "SPY",
                          color = "cornflowerblue") %>%
            hc_navigator(enabled = FALSE) %>%
            hc_scrollbar(enabled = FALSE) %>%
            hc_exporting(enabled = TRUE) %>%
            hc_legend(enabled = TRUE,
                      align = "right",
                      verticalAlign = "middle",
                      layout = "vertical")})

    # value boxes
    output$box_portfolio <- renderValueBox({
        valueBox(value = tags$p(round(portfolio_sharpe(), 4),
                                style = "font-size: 70%;"),
                 color = "blue",
                 subtitle = "Sharpe Ratio of Portfolio")})

    output$box_market <- renderValueBox({
        valueBox(value = tags$p(round(market_sharpe(), 4),
                                style = "font-size: 70%;"),
                 color = "green",
                 subtitle = "Sharpe Ratio of S&P 500")})
}

```

```{r}
# shinyApp(ui, server)
```

# Capital Asset Pricing Model

The CAPM is a model created by William Sharpe that estimates the return of an asset based on the return of the market and an asset's linear relationship to the return of the market. That linear relationship is the stock's beta coefficient. Beta can be thought of as the stock's sensitivity to the market, or its riskiness with respect to the market.

Fama and French have written that CAPM "is the centerpiece of MBA investment courses. Indeed, it is often the only asset pricing model taught in these courses... unfortunately, the empirical record of the model is poor".

This is still useful because it can serve as a nice template for more complex models.

# CAPM and Market Returns

Our first step is to make a choice about which asset to use as a proxy for the market return. We will go with SPY ETF, effectively treating S&P500 as the market. This makes our calculations uninteresting because:

1. S&P is 25% of our portfolio 
2. We have chosen assets and a time period (2013 - ) in which correlations with SPY have been high

```{r}
# import prices for SPY, calculate monthly returns
market_returns_xts <- get_returns("SPY", "SPY", "2012-12-31")

portfolio_returns_xts_rebalanced_monthly %>% head(2)
market_returns_xts %>% head(2)

market_returns_tidy <- market_returns_xts %>%
    tk_tbl(preserve_index = TRUE,
           rename_index = "date") %>%
    na.omit() %>%
    select(date, returns = SPY)

# since we are regressing on market returns
# ensure that num portfolio returns == market returns
portfolio_returns_dplyr %>% 
    mutate(market_returns = market_returns_tidy$returns) %>%
    head()
```

# Calculating CAPM Beta

Portfolio beta is equal to the covariance of the portfolio returns and market returns, divided by the variance of the market returns

$\beta_{portfolio} = \mathrm{cov}(R_p, R_m) / \sigma_m$

Our portfolio beta is equal to

```{r}
cov(portfolio_returns_xts_rebalanced_monthly,
    market_returns_tidy$returns) /
    var(market_returns_tidy$returns)
```

The beta is near 1, which is not surprising since SPY is a big part of the portfolio. 

We can also find the beta by finding the beta of each of our assets and then multiplying by asset weights.

$\beta_{portfolio} = \sum_{i = 1}^n W_i \beta_i$

We need to regress each of our individual asset returns on the market return. We can use a combination of lm and map.

```{r}
# nest by asset
beta_assets <- asset_returns_long %>%
    nest(-asset)

# run a model on each asset
beta_assets %<>%
    mutate(model = map(data, ~ lm(returns ~ market_returns_tidy$returns,
                                  data = .)))

# get results
beta_assets %<>%
    mutate(model = map(model, broom::tidy))

# unnest
beta_assets %<>%
    unnest(model) %>%
    mutate_if(is.numeric, ~ round(.x, 4))

# delete the intercepts to just have betas
beta_assets %<>%
    filter(term != "(Intercept)") %>%
    select(-term)
```

We see that SPY has a beta of 1 with itself. This is a good sanity check.

```{r}
weights %>%
    map2(., 1:5, ~ .x * beta_assets$estimate[.y]) %>%
    reduce(sum)
```

# Calculating CAPM Beta in the xts World

```{r}
(beta_builtin_xts <- CAPM.beta(portfolio_returns_xts_rebalanced_monthly,
                              market_returns_xts))
```

# Calculating CAPM Beta in the Tidyverse

```{r}
(beta_dplyr <- portfolio_tq %>%
    ## rowwise() %>% #
    do(model = lm(returns ~ market_returns_tidy$returns,
                  data = .)) %>%
    tidy(model) %>%
    mutate(term = c("alpha", "beta")) %>%
    select(estimate) %>%
    slice(2))
```

# Calculate CAPM Beta in Tidyquant

```{r}
beta_tq <- portfolio_tq %>%
    mutate(market_return = market_returns_tidy$returns) %>%
    na.omit() %>%
    tq_performance(Ra = returns,
                   Rb = market_return,
                   performance_fun = CAPM.beta) %>%
    colnamer("beta_tq")
```

# Visualizing CAPM with ggplot

```{r}
portfolio_tq %>%
    mutate(market_returns = market_returns_tidy$returns) %>%
    ggplot(aes(x = market_returns, y = returns)) +
    geom_point(color = "cornflowerblue") +
    geom_smooth(method = "lm", color = "mediumpurple", se = FALSE, lty = 2, alpha = 0.5) +
    ylab("Portfolio Returns") + xlab("Market Returns")
```

# Augmenting Our Data

```{r}
portfolio_augmented <- portfolio_tq %>%
    do(model = lm(returns ~ market_returns_tidy$returns, data = .)) %>%
    augment(model) %>%
    rename(mkt_rtns = market_returns_tidy.returns) %>%
    select(returns, mkt_rtns, .fitted) %>%
    mutate(date = portfolio_tq$date)
```

Lets view how well our fitted return values match the actual return values

```{r}
portfolio_augmented %>%
    select(date, returns, .fitted) %>%
    gather(type, data, -date) %>%
    ggplot(aes(x = date, y = data, color = type)) +
    geom_line() +
    xlab("date")
```

# Visualizing CAPM with Highcharter

```{r}
highchart() %>%
    hc_title(text = "Portfolio vs. Market Returns Scatter") %>%
    hc_add_series(portfolio_augmented,
                  type = "scatter",
                  color = "cornflowerblue",
                  hcaes(x = round(mkt_rtns, 4),
                        y = round(returns, 4)),
                  name = "Returns") %>%
    hc_xAxis(title = list(text = "Market Returns")) %>%
    hc_yAxis(title = list(text = "Portfolio Returns")) %>%
    hc_add_theme(hc_theme_flat()) %>%
    hc_exporting(enabled = TRUE)  
```

The plot looks good, but we want to add a tool tip to show the date on hover

```{r}
tt_js <- "function() {
    return('port return: ' + this.y +
           '<br> mkt return: ' + this.x +
           '<br> date: ' + this.point.date)}" %>% str_squish()

highchart() %>%
    hc_title(text = "Scatter Plot with Date") %>%
    hc_add_series(portfolio_augmented,
                  type = "scatter",
                  color = "cornflowerblue",
                  hcaes(x = round(mkt_rtns, 4),
                        y = round(returns, 4),
                        date = date),
                  name = "Returns") %>%
    hc_add_series(portfolio_augmented,
                  type = "line",
                  hcaes(x = mkt_rtns, y = .fitted),
                  name = "CAPM Beta = Regression Slope") %>% 
    hc_tooltip(formatter = JS(tt_js)) %>% 
    hc_xAxis(title = list(text = "Market Returns")) %>%
    hc_yAxis(title = list(text = "Portfolio Returns")) %>%
    hc_add_theme(hc_theme_flat()) %>%
    hc_exporting(enabled = TRUE) 
```

# Shiny App CAPM

```{r}
symbols <- c("SPY", "EFA", "IJS", "EEM", "AGG")

make_input <- function(symbol, index) {
    fluidRow(
        column(6,
               textInput(paste0("stock_", symbol),
                         paste("Stock", index),
                         symbol)),
        column(6,
               numericInput(paste0("wt_", index),
                            "Portfolio %",
                            value = 20,
                            min = 0,
                            max = 100)))
}

stock_inputs <- imap(symbols, ~ make_input(.x, .y))

date_input <- fluidRow(
    column(6,
           dateInput("date",
                     "Starting Date",
                     "2013-01-01",
                     format = "yyyy-mm-dd")))

buttons <- fluidRow(
    column(6,
           actionButton("go",
                        "Submit",
                        class = "btn btn-success")))

ui <- fluidPage(
    theme = shinythemes::shinytheme("paper"),
    sidebarLayout(
        sidebarPanel(
            !!!stock_inputs,
            date_input,
            br(),
            buttons),
        mainPanel(
            fluidRow(renderHighchart("hchart_plot")),
            fluidRow(renderTable("results")))))


```

```{r}
get_prices <- function(syms, from_date) {
    getSymbols(syms,
               src = "yahoo",
               from = from_date,
               auto.assign = TRUE,
               warnings = FALSE) %>%
        map(., ~ Ad(get(.x))) %>%
        reduce(merge) %>%
        `colnames<-`(syms)
}

get_tq_returns <- function(data, weights) {
    data %>%
            to.monthly(indexAt = "last",
                       OHLC = FALSE) %>%
        tk_tbl(preserve_index = TRUE,
               rename_index = "date") %>%
        slice(-1) %>%
        gather(asset, returns, -date) %>%
        group_by(asset) %>%
        mutate(returns = (log(returns) - log(lag(returns)))) %>% 
        tq_portfolio(assets_col = asset,
                     returns_col = returns,
                     weights = weights,
                     col_rename = "returns")
}

server <- function(input, output, session) {
    # get data
    prices <- eventReactive(input$go, {
        # place parameters in vectors
        syms <- c(input$stock_SPY, input$stock_EFA,
                  input$stock_IJS, input$stock_EEM,
                  input$stock_AGG)

        weights <- c(input$wt_1 / 100, input$wt_2 / 100,
                     input$wt_3 / 100, input$wt_4 / 100,
                     input$wt_5 / 100)

        # get prices of the symbols
        get_prices(syms, input$date)})

    portfolio_returns <- eventReactive(input$go, {
        prices() %>% get_tq_returns(weights = weights)})

    market_returns <- eventReactive(input$go, {
        get_prices(syms = "SPY",
                   from_date = input$date) %>%
            get_tq_returns(weights = 1)})

    # transform data
    beta <- eventReactive(input$go, {
        portfolio_returns() %>%
            do(model = lm(returns ~ market_returns()[["returns"]],
                          data = .)) %>%
            augment(model) %>%
            rename(mkt_rtns = market_return.returns) %>%
            select(returns, mkt_rtns, .fitteed) %>%
            mutate(date = portfolio_returns()[["date"]])})

    # plot data
    output$hchart_plot <- renderHighchart({
        highchart() %>%
            hc_title(text = "Scatter with Regression Line") %>%
            hc_add_series(beta(),
                          type = "scatter",
                          color = "cornflowerblue",
                          hcaes(x = round(mkt_rtns, 4),
                                y = round(returns, 4),
                                date = date),
                          name = "Returns") %>%
            hc_add_series(beta(),
                          type = "line",
                          enableMouseTracking = FALSE,
                          hcaes(x = mkt_rtns, y = .fitted),
                          name = "CAPM Beta = Slope of Line") %>%
            hc_xAxis(title = list(text = "Market Returns")) %>%
            hc_yAxis(title = list(text = "Portfolio Returns")) %>%
            hc_add_theme(hc_theme_flat()) %>%
            hc_exporting(enabled = TRUE)})

    # plot table
    output$results <- renderTable({
        portfolio_returns() %>%
            do(model = lm(returns ~ market_returns()[["returns"]],
                          data = .)) %>%
            tidy(model) %>%
            mutate(term = c("alpha", "beta"))},
        digits = 4)
}
```

```{r}
shinyApp(ui, server)
```
