# Building a Portfolio

```{r}
library(tidyverse)
library(shiny)
library(xts)
library(quantmod)
library(lubridate)
library(readxl)
library(highcharter)
library(tidyquant)
library(timetk)
library(tibbletime)
library(quantmod)
library(PerformanceAnalytics)
```

Now we collect individual returns into a portfolio, which is a weighted set of asset returns. Accordingly, the first thing we need to do is to assign a weight to each asset. 

```{r}
# set weights
weights <- c(0.25, 0.25, 0.2, 0.2, 0.1)

# combine in df
tibble(weights, symbols)

# sanity check
tibble(weights, symbols) %>%
    summarise(total_weight = sum(weights))
```

Now we can use those weights to convert the weights of 5 assets into the returns of 1 portfolio. The return of a multi asset portfolio is the sum of the weighted returns of each asset. 

# Portfolio Returns in the xts World 

```{r}
Return.portfolio(asset_returns_xts,
                 weights = weights,
                 rebalance_on = "months") %>%
    `colnames<-`("returns") -> portfolio_returns_xts_rebalanced_monthly
```

# Portfolio Returns in the Tidyverse

```{r}
weight_table <- tibble(weights, symbols)

asset_returns_long %>%
    group_by(asset) %>%
    left_join(weight_table, by = c("asset" = "symbols")) %>%
    mutate(weighted_returns = returns * weights) %>%
    group_by(date) %>%
    summarise(returns = sum(weighted_returns)) -> portfolio_returns_dplyr
```

# Portfolio Returns in the tidyquant World 

```{r}
asset_returns_long %>%
    tq_portfolio(assets_col = asset,
                 returns_col = returns,
                 weights = weights,
                 col_rename = "returns",
                 rebalance_on = "months") -> portfolio_tq
```

# Visualizing Portfolio Returns in the xts World 

```{r}
highchart(type = "stock") %>%
    hc_title(text = "Portfolio Monthly Returns") %>%
    hc_add_series(portfolio_returns_xts_rebalanced_monthly$returns,
                  name = "Rebalanced Monthly",
                  color = "mediumpurple") %>%
    hc_add_theme(hc_theme_economist()) %>%
    hc_navigator(enabled = FALSE) %>%
    hc_scrollbar(enabled = FALSE) %>%
    hc_legend(enabled = FALSE) %>%
    hc_exporting(enabled = TRUE)
```

```{r}
hchart(hist(portfolio_returns_xts_rebalanced_monthly$returns,
            breaks = 50,
            plot = FALSE),
       color = "forestgreen",
       name = "Portfolio") %>%
    hc_title(text = "Portfolio Returns Distribution") %>%
    hc_add_theme(hc_theme_flat()) %>%
    hc_exporting(enabled = TRUE)
```

# Visualizing Portfolio Returns in the Tidyverse

## Scatterplot

```{r}
portfolio_tq %>%
    ggplot(aes(x = date, y = returns)) +
    geom_point(color = "mediumpurple") +
    xlab("date") + ylab("monthly return") +
    theme_update(plot.title = element_text(hjust = 0.5)) +
    ggtitle("Portfolio Returns Scatter")
```

## Histogram

```{r}
portfolio_tq %>%
    ggplot(aes(x = returns)) +
    geom_histogram(binwidth = .005,
                   fill = "mediumpurple",
                   color = "black") +
    ggtitle("Portfolio Returns Distribution") +
    geom_vline(xintercept = 0, lty = 2) +
    theme_update(plot.title = element_text(hjust = 0.5))
```

We can also compare the portfolio distribution to those of our individual assets by layering on geoms. 

```{r}
asset_returns_long %>%
    ggplot(aes(x = returns, fill = asset)) +
    geom_histogram(alpha = 0.15, binwidth = .01) +
    geom_histogram(data = portfolio_tq,
                   fill = "mediumpurple",
                   color = "black",
                   binwidth = .01,
                   alpha = 0.4) +
    ggtitle("Portfolio and Asset Monthly Returns") +
    theme_update(plot.title = element_text(hjust = 0.5))
```

```{r}
portfolio_tq %>%
    ggplot(aes(x = returns)) +
    geom_histogram(binwidth = .01,
                   fill = "mediumpurple",
                   color = "black") +
    geom_density(alpha = 0.5, color = "blue", lty = 2) +
    xlab("monthly returns") +
    ylab("distribution") +
    theme_update(plot.title = element_text(hjust = 0.5)) +
    ggtitle("Portfolio Histogram and Density")
```

# Shiny App Portfolio Returns 

We want to empower an end user to do the following:

1. Build a portfolio by choosing assets and weights 
2. choose a start date
3. choose a rebalancing frequency
4. calculate portfolio returns 
5. visualize the portfolio returns on histogram and density charts 

```{r}

symbols <- c("SPY", "EFA", "IJS", "EEM", "AGG")

make_input <- function(symbol, index) {
    fluidRow(
        column(6,
               textInput(paste0("stock_", symbol),
                         paste("Stock", index),
                         symbol)),
        column(5,
               numericInput(paste0("wt_", index),
                            "Portfolio %",
                            value = 20,
                            min = 0,
                            max = 100)))
}

stock_inputs <- imap(symbols, ~ make_input(.x, .y))

date_input <- fluidRow(
    column(7,
           dateInput("date",
                     "Starting Date",
                     "2013-01-01",
                     format = "yyyy-mm-dd")))

rebal_input <- fluidRow(
    column(7,
           selectInput("rebal",
                       "Rebalance Frequency",
                       choices = c("Yearly" = "years",
                                   "Monthly" = "months",
                                   "Weekly" = "weeks"))))

ui <- fluidPage(
    theme = shinythemes::shinytheme("paper"),
    sidebarLayout(
        sidebarPanel(!!!stock_inputs,
                     date_input,
                     rebal_input,
                     actionButton("go", "Submit")),
        mainPanel(
            fluidRow(
                tabsetPanel(
                    tabPanel("Histogram",
                             plotOutput("plot_1")),
                    tabPanel("Density",
                             plotOutput("plot_2")),
                    tabPanel("Both",
                             plotOutput("plot_3")))))))
```



```{r}
get_returns <- function(syms, weights, from_date, rebal) {
    # pull in info from yahoo

    prices <- getSymbols(syms,
               src = "yahoo",
               ## from = from_date,
               auto.assign = TRUE,
               warnings = FALSE) %>%
        map(., ~ Ad(get(.x))) %>%
        reduce(merge) %>%
        `colnames<-`(syms)
    
    # convert to monthly gains
    asset_returns <- prices %>% to.monthly(indexAt = "last",
                                           OHLC = FALSE) %>%
        tk_tbl(preserve_index = TRUE,
               rename_index = "date") %>%
        gather(asset, returns, -date) %>%
        group_by(asset) %>%
        mutate(returns = (log(returns) - log(lag(returns))))

     # check portfolio returns
    (portfolio_returns <- tq_portfolio(data = asset_returns,
                                       assets_col = asset,
                                       returns_col = returns,
                                       weights = weights,
                                       rebalance_on = rebal,
                                       col_rename = "returns"))
}

get_returns(symbols, c(.20, .20, .20, .20, .20), "2018-01-02", "weeks")

server <- function(input, output, session) {
   portfolio_returns <- eventReactive(input$go, {
       syms <- c(input$stock_SPY, input$stock_EFA,
                 input$stock_IJS, input$stock_EEM,
                 input$stock_AGG)

       weights <- c(input$wt_1 / 100, input$wt_2 / 100,
                    input$wt_3 / 100, input$wt_4 / 100,
                    input$wt_5 / 100)

       get_returns(syms = syms,
                   weights = weights,
                   from_date = input$date,
                   rebal = input$rebal)
       })

   output$plot_1 <- renderPlot({
                             portfolio_returns() %>%
                                 ggplot(aes(x = returns)) +
                                 geom_histogram(alpha = 0.25,
                                                binwidth = .01,
                                                fill = "blue")})

   output$plot_2 <- renderPlot({
                             portfolio_returns() %>%
                                 ggplot(aes(x = returns)) +
                                 geom_density(size = 1,
                                              color = "blue",
                                              lty = 2)})

   output$plot_3 <- renderPlot({
       portfolio_returns() %>%
           ggplot(aes(x = returns)) +
           geom_histogram(alpha = 0.25,
                          binwidth = .01,
                          fill = "blue") +
           geom_density(size = 1,
                        alpha = 0.5,
                        color = "forestgreen",
                        lty = 3)})
}

# shinyApp(ui, server)
```
