# Practice and Applications

```{r}
library(tidyverse)
library(magrittr)
library(shiny)
```

# Component Contribution to Standard Deviation

Our goal in this chapter is to investigate how each of our five assets contributes to portfolio standard deviation. 

From a substantive perspective, we want to ensure that our risk has not gotten too concentrated in one asset. This could lead to a less diversified portfolio than intended.

The percentage contribution to portfolio standard deviation of an asset is defined as 

(marginal contribution of the asset * weight of the asset) / (portfolio standard deviation)

```{r}
# calculate portfolio standard deviation
covariance_matrix <- cov(asset_returns_xts)
wts <- rep(0.2, 5)

sd_portfolio <- sqrt(t(wts) %*% covariance_matrix %*% wts) %>% .[1, 1]

# find marginal contribution of each asset
marginal_contribution <- wts %*% covariance_matrix / sd_portfolio

# get component contributions by weighting marginal contributions
component_contribution <- marginal_contribution * wts

# check results sum to our portfolio SD
rowSums(component_contribution)

# get percentage contribution of each asset
component_percentages <- component_contribution / sd_portfolio

component_percentages %>% round(3)
```

# Component Contribution with a function

```{r}
comp_cont <- function(asset_returns, weights) {
    covariance_matrix <- asset_returns %>% cov()
    sd_portfolio <- sqrt(t(weights) %*% covariance_matrix %*% weights) %>% .[1, 1]
    marginal_contribution <- weights %*% covariance_matrix / sd_portfolio
    component_contribution <- marginal_contribution * weights
    component_percentages <- component_contribution / sd_portfolio
    
    component_percentages %>%
        as_tibble() %>%
        gather(asset, contribution)
}

asset_returns_xts %>% comp_cont(weights)
```

```{r}
asset_returns_dplyr %>%
    select(-date) %>%
    comp_cont(weights) -> percentages_tibble
```

# Visualizing Component Contribution

```{r}
percentages_tibble %>%
    ggplot(aes(x = asset, y = contribution)) +
    geom_col(fill = "mediumpurple", color = "black", width = 0.6) +
    scale_y_continuous(labels = scales::percent) + 
    ggtitle("% Contribution to Standard Deviation") +
    labs(x = "Asset", y = "% Contribution to Risk")
```

It would also be useful to see a chart that compares asset weight to risk contribution.

```{r}
percentages_tibble %>%
    mutate(weights = weights) %>%
    gather(type, percent, -asset) %>%
    group_by(type) %>%
    ggplot(aes(x = asset, y = percent, fill = type)) +
    geom_col(position = "dodge") +
    scale_y_continuous(labels = scales::percent) +
    ggtitle("% Contribution to Volatility")
```

From this plot we see that AGG, the bond fund, has done a good job as a volatility dampener. It has a 10% allocation, but contributes almost 0 to volatility.

# Rolling Component Contribution

The previous section showed us the total contribution of each asset over the life of the portfolio, but now we wish to understand risk components over time.

```{r}
interval_sd <- function(asset_returns, weights, start_date, window) {
	# create dates
    start_dte <- asset_returns$date[start_date]
    end_date <- asset_returns$date[c(start_date + window)]

    asset_returns %<>%
        filter(date >= start_dte &
               date < end_date) %>%
        select(-date)

    comp_cont(asset_returns, weights) %>%
        mutate(date = ymd(end_date)) %>%
        select(date, everything()) %>%
        spread(asset, contribution) %>%
        mutate_if(is.numeric, ~ .x * 100)
}

interval_sd(asset_returns_dplyr, weights, 1, 24)
```

This function returns the contribution to risk for each asset from January 2013 through January 2015.

```{r}
interval_sd(asset_returns_dplyr, weights, start_date = 2, window = 24)
```

```{r}
map_df(seq_len(nrow(asset_returns_dplyr) - window),
       ~ interval_sd(asset_returns_dplyr,
                     weights,
                     .x,
                     window)) -> portfolio_vol_comps

portfolio_vol_comps %>% tail()
```

# Visualizing Rolling Component Contribution

Line Plot

```{r}
portfolio_vol_comps %>%
    gather(asset, contribution, -date) %>%
    group_by(asset) %>%
    ggplot(aes(x = date)) +
    geom_line(aes(y = contribution, color = asset)) +
    scale_y_continuous(labels = function(x) paste0(x, "%"))
```

Stacked Plot

```{r}
portfolio_vol_comps %>%
    gather(asset, contribution, -date) %>%
    group_by(asset) %>%
    ggplot(aes(x = date, y = contribution)) +
    geom_area(aes(fill = asset), color = "black") +
    scale_y_continuous(labels = function(x) paste0(x, "%"))
```

In Highcharter

```{r}
portfolio_vol_comps %>%
    tk_xts(date_var = date,
           silent = TRUE) -> port_comp_xts

highchart(type = "stock") %>%
    hc_title(text = "Volatility Contribution") %>%
    hc_add_series(port_comp_xts[, 1], name = symbols[5]) %>%
    hc_add_series(port_comp_xts[, 2], name = symbols[4]) %>%
    hc_add_series(port_comp_xts[, 3], name = symbols[2]) %>%
    hc_add_series(port_comp_xts[, 4], name = symbols[3]) %>%
    hc_add_series(port_comp_xts[, 5], name = symbols[1]) %>%
    hc_yAxis(labels = list(format = "{value}%"),
             max = max(port_comp_xts) + 5,
             min = min(port_comp_xts) - 5,
             opposite = FALSE) %>%
    hc_navigator(enabled = FALSE) %>%
    hc_scrollbar(enabled = FALSE) %>%
    hc_add_theme(hc_theme_flat()) %>%
    hc_legend(enabled = TRUE)

symbols
port_comp_xts %>% head(1)
```

```{r}
highchart() %>%
    hc_chart(type = "area") %>%
    hc_title(text = "Volatility Contribution") %>%
    hc_plotOptions(area = list(
                       stacking = "percent",
                       linrColor = "#ffffff",
                       lineWidth = 1,
                       marker = list(
                           lineWidth = 1,
                           lineColor = "#ffffff"))) %>%
    hc_add_series(port_comp_xts[, 1], name = symbols[5]) %>%
    hc_add_series(port_comp_xts[, 2], name = symbols[4]) %>%
    hc_add_series(port_comp_xts[, 3], name = symbols[2]) %>%
    hc_add_series(port_comp_xts[, 4], name = symbols[3]) %>%
    hc_add_series(port_comp_xts[, 5], name = symbols[1]) %>%
    hc_yAxis(labels = list(format = "{value}%"),
             opposite = FALSE) %>%
    hc_xAxis(type = "datetime") %>%
    hc_tooltip(pointFormat = "<span style=\"color:{series.color}\">
          {series.name}</span>:<b>{point.percentage:.1f}%</b></br>",
          shared = TRUE) %>%
    hc_navigator(enabled = FALSE) %>%
    hc_scrollbar(enabled = FALSE) %>%
    hc_add_theme(hc_theme_flat()) %>%
    hc_legend(enabled = TRUE)
```

# Shiny App Component Contribution

```{r}
symbols <- c("SPY", "EFA", "IJS", "EEM", "AGG")

make_input <- function(symbol, index) {
    fluidRow(
        column(6,
               textInput(paste0("stock_", symbol),
                         paste("Stock", index),
                         symbol)),
        column(6,
               numericInput(paste0("wt_", index),
                            "Portfolio %",
                            value = 20,
                            min = 0,
                            max = 100)))
}

stock_inputs <- imap(symbols, ~ make_input(.x, .y))

date_input <- fluidRow(
    column(6,
           dateInput("date",
                     "Starting Date",
                     "2013-01-01",
                     format = "yyyy-mm-dd")),
    column(6,
           numericInput("window",
                        "Window",
                        min = 1,
                        max = 24,
                        value = 6)))

buttons <- actionButton("go",
                        "Submit",
                        class = "btn btn-success")

ui <- fluidPage(
    theme = shinythemes::shinytheme("simplex"),
    sidebarLayout(
    sidebarPanel(!!!stock_inputs,
                 date_input,
                 buttons,
                 width = 2),
    mainPanel(
        fluidRow(tabPanel("Asset Contribution",
                          highchartOutput("hchart_contrib")),
                 tabPanel("Stacked Area Chart",
                          highchartOutput("hchart_stacked"))),
              fluidRow("barchart",
                       renderPlot("gg_out")))))
```

```{r}
server <- function(input, output, session) {
    # calculate asset returns
    asset_returns_dplyr <- eventReactive(input$go, {
        syms <- c(input$stock_SPY, input$stock_EFA,
                  input$stock_IJS, input$stock_EEM,
                  input$stock_AGG)
        
        prices <- getSymbols(syms,
                             src = "yahoo",
                             from = input$date,
                             auto.assign = TRUE,
                             warnings = FALSE) %>%
            map(., ~ Ad(get(.x))) %>%
            reduce(merge) %>%
            colnamer(syms)

        asset_returns_dplyr <- prices %>%
            to.monthly(indexAt = "last", OHLC = FALSE) %>%
            tk_tbl(preserve_index = TRUE,
                   rename_index = "date") %>%
            gather(asset, returns, -date) %>%
            group_by(asset) %>%
            mutate(returns = (log(returns) - log(lag(returns)))) %>%
            spread(asset, returns) %>%
            select(date, symbols) %>%
            slice(-1)})

    # calculate overall contribution to risk
    percentages_tibble <- eventReactive(input$go, {
        weights <- c(input$wt_1 / 100, input$wt_2 / 100,
                     input$wt_3 / 100, input$wt_4 / 100,
                     input$wt_5 / 100)

        asset_returns_xts <- asset_returns_dplyr() %>%
            tk_xts(date_col = date)

        port_vol_contr <- StdDev(asset_returns_xts,
                                 weights = weights,
                                 portfolio_method = "component")

        percentages_tibble <- port_vol_contr$pct_contrib_StdDev %>%
            tk_tbl(preserve_index = FALSE) %>%
            mutate(asset = symbols) %>%
            rename("risk_contribution" = data) %>%
            mutate(risk_contribution = round(risk_contribution, 4) * 100,
                   weights = weights * 100) %>%
            select(asset, everything())})

    # calculate rolling contribution to stddev
    portfolio_vol_comps_xts <- eventReactive(input$go, {
        asset_returns_dplyr <- asset_returns_dplyr()

        map_df(1:(nrow(asset_returns_dplyr) - input$window),
               interval_sd(asset_returns = asset_returns_dplyr,
                           weights = weights,
                           start_date = 1,
                           window = input$window)) %>%
            tk_xts(date_col = date) -> portfolio_vol_comps_xts})

    output$hchart_contrib <- renderHighchart({
        port_vol_comps <- portfolio_vol_comps_xts()

        highchart(type = "stock") %>%
            hc_title(text = "Volatility Contribution") %>%
            hc_add_series(port_vol_comps[, 1],
                          name = names(port_vol_comps[, 1])) %>%
            hc_add_series(port_vol_comps[, 2],
                          name = names(port_vol_comps[, 2])) %>%
            hc_add_series(port_vol_comps[, 3],
                          name = names(port_vol_comps[, 3])) %>%
            hc_add_series(port_vol_comps[, 4],
                          name = names(port_vol_comps[, 4])) %>%
            hc_add_series(port_vol_comps[, 5],
                          name = names(port_vol_comps[, 5])) %>%
            hc_add_theme(hc_theme_flat()) %>%
            hc_yAxis(labels = list(format = "{value}%"),
                     opposite = FALSE,
                     min = min(port_vol_comps) - 5,
                     max = max(port_vol_comps) + 5) %>%
            hc_navigator(enabled = FALSE) %>%
            hc_scrollbar(enabled = FALSE)})


    output$hchart_stacked <- renderHighchart({
        port_vol_comps <- portfolio_vol_comps_xts()

        highchart() %>%
            hc_chart(type = "area") %>%
            hc_title(text = "Stacked Volatility Contribution") %>%
            hc_plotOptions(area = list(
                               stacking = "percent",
                               lineColor = "#ffffff",
                               lineWidth = 1,
                               marker = list(
                                   lineWidth = 1,
                                   lineColor = "#ffffff"))) %>% 
            hc_add_series(port_vol_comps[, 1],
                          name = names(port_vol_comps[, 1])) %>%
            hc_add_series(port_vol_comps[, 2],
                          name = names(port_vol_comps[, 2])) %>%
            hc_add_series(port_vol_comps[, 3],
                          name = names(port_vol_comps[, 3])) %>%
            hc_add_series(port_vol_comps[, 4],
                          name = names(port_vol_comps[, 4])) %>%
            hc_add_series(port_vol_comps[, 5],
                          name = names(port_vol_comps[, 5])) %>%
            hc_add_theme(hc_theme_flat()) %>%
            hc_yAxis(labels = list(format = "{value}%"),
                     opposite = FALSE) %>%
            hc_xAxis(type = "datetime") %>% 
            hc_navigator(enabled = FALSE) %>%
            hc_scrollbar(enabled = FALSE) %>%
            hc_legend(enabled = TRUE)}) 

    output$gg_out <- renderPlot({
        percentages_tibble() %>%
            gather(type, percent, -asset) %>%
            group_by(type) %>%
            mutate(percent = percent / 100) %>%
            ggplot(aes(x = asset, y = percent, fill = type)) +
            geom_col(position = "dodge") +
            scale_y_continuous(labels = scales::percent) +
            ggtitle("Percent Contribution to Volatility")})
}

```

```{r}
# shinyApp(ui, server)
```

# Monte Carlo Simulation

## Simulating Growth of a Dollar

```{r}
mean_port_return <- mean(portfolio_tq$returns)
stddev_port_return <- sd(portfolio_tq$returns)
```

Now we can use the rnorm function to create a distribution with mean and sd equal to our portfolio.

```{r}
simulated_monthly_returns <- rnorm(120,
                                   mean_port_return,
                                   stddev_port_return)
```

Next we calculate how a dollar would have grown given those random monthly returns. We first add a 1 to each of our monthly returns, since we start with $1

```{r}
(sim_returns_add_1 <- tibble(c(1, 1 + simulated_monthly_returns)) %>%
    colnamer("returns"))
```

This data can now be converted into the cumulative growth of a dollar. 

```{r}
(simulated_growth <- sim_returns_add_1 %>%
    mutate(growth = accumulate(returns, `*`)))
```

Are these values reasonable? What compound annual growth rate (CAGR) is implied by this simulation?

```{r}
(cagr <- ((simulated_growth$growth[nrow(simulated_growth)] ^ (1 / 10)) - 1) * 100)
```

# Several Simulation Functions

```{r}
sim_accum <- function(init, N, mean, stddev) {
    tibble(c(init, 1 + rnorm(N, mean, stddev))) %>%
        colnamer("returns") %>%
        mutate(growth = accumulate(returns, `*`)) %>%
        select(growth)}

sim_accum(1, 120, mean_port_return, stddev_port_return)
```

# Running Multiple Simulations

Suppose we wish to run 50 simulations to get a feel for how the randomness is distributed. 

```{r}
sims <- 50
(starts <- rep(1, sims) %>%
    set_names(paste0("sim_", 1:sims)))
```

We will store the results of our simulations in starts.

We want to apply our simulation function to each of the 50 columns of the starts matrix. 

map_dfc() takes a vector and applies a function to it. In this case, each column of starts.

```{r}
monte_carlo_sim_50 <- map_dfc(starts, ~ sim_accum(.x, 120,
                                                  mean_port_return,
                                                  stddev_port_return))

monte_carlo_sim_50 %>% tail()
```

We now have 50 simulations of the growth of the dollar and we simulated that growth over 120 months. To visualize this, we still need a month column. 

```{r}
monte_carlo_sim_50 %>%
    mutate(month = seq(1:nrow(.))) %>%
    select(month, everything()) %>%
    colnamer(c("month", names(starts))) %>%
    mutate_all(~ round(.x, 2))
```

An alternative method is to use the rerun() function from purrr. 

```{r}
monte_carlo_rerun <- rerun(5, sim_accum(1, 120,
                                        mean_port_return,
                                        stddev_port_return))
```

We now have 5 dataframes in a list. We can consolidate that list into one tibble.

```{r}
set.seed(8888)

reruns <- 50

monte_carlo_rerun <- rerun(reruns,
                           sim_accum(1, 120,
                                     mean_port_return,
                                     stddev_port_return)) %>%
    simplify_all() %>%
    `names<-`(paste0("sim_", 1:reruns)) %>%
    as_tibble() %>%
    mutate(month = seq(1:nrow(.))) %>%
    select(month, everything())
```

# Visualizing Simulations with ggplot

```{r}
monte_carlo_rerun %>%
    gather(sim, growth, -month) %>%
    group_by(sim) %>%
    ggplot(aes(x = month, y = growth, color = sim)) +
    geom_line() +
    theme(legend.position = "none")
```

We can check the min, max, and median simulations with summarize

```{r}
(sim_sum <- monte_carlo_rerun %>%
    gather(sim, growth, -month) %>%
    group_by(sim) %>%
     summarize(final = last(growth)) %>%
     summarise(max = max(final),
               min = min(final),
               median = median(final)))
```

```{r}
monte_carlo_rerun %>%
    gather(sim, growth, -month) %>%
    group_by(sim) %>%    
    filter(last(growth) == sim_sum$max ||
           last(growth) == sim_sum$median ||
           last(growth) == sim_sum$min) %>%
    ggplot(aes(x = month, y = growth)) +
    geom_line(aes(color = sim))
```

Since we have a large range of possible end values, let's examine the quantiles for those values

```{r}
probs <- c(.005, .025, .25, .5, .75, .975, .995)

sim_final_quantile <- monte_carlo_rerun %>%
    gather(sim, growth, -month) %>%
    group_by(sim) %>%
     summarise(final = last(growth))

(quantiles <- quantile(sim_final_quantile$final, probs = probs) %>%
    tibble() %>%
    colnamer("value") %>%
    mutate(probs = probs) %>%
    spread(probs, value))
```

# Visualizing Simulations with Highcharter

```{r}
mc_gathered <- monte_carlo_rerun %>%
    gather(sim, growth, -month) %>%
    group_by(sim)
```

```{r}
mc_gathered %>%
    hchart(type = "line",
           hcaes(y = growth,
                 x = month,
                 group = sim)) %>%
    hc_title(text = "50 Simulations") %>%
    hc_xAxis(title = list(text = "months")) %>%
    hc_yAxis(title = list(text = "dollar growth"),
             labels = list(format = "${value}")) %>%
    hc_add_theme(hc_theme_flat()) %>%
    hc_exporting(enabled = TRUE) %>%
    hc_legend(enabled = FALSE)
```

```{r}
mc_gathered %>%
    filter(last(growth) == sim_sum$max ||
           last(growth) == sim_sum$median ||
           last(growth) == sim_sum$min) %>%
    group_by(sim) %>%
    hchart(type = "line",
           hcaes(y = growth,
                 x = month,
                 group = sim)) %>%
    hc_title(text = "Min, Max, Median Simulations") %>%
    hc_xAxis(title = list(text = "months")) %>%
    hc_yAxis(title = list(text = "dollar growth"),
             labels = list(format = "${value}")) %>%
    hc_add_theme(hc_theme_flat()) %>%
    hc_exporting(enabled = TRUE) %>%
    hc_legend(enabled = FALSE)
```

# Shiny App Monte Carlo

```{r}

```
