# Risk 

```{r}
library(magrittr)
library(tidyverse)
library(shiny)
library(xts)
library(quantmod)
library(lubridate)
library(readxl)
library(highcharter)
library(tidyquant)
library(timetk)
library(tibbletime)
library(quantmod)
library(PerformanceAnalytics)
```

This section on risk will focus on calculating the standard deviation, skewness, and kurtosis of portfolio returns. 

We will mainly focus on standard deviation  as it measures the extent to which a portfolios returns are dispersed around their mean. If returns are more dispersed, the portfolio has a higher standard deviation and is seen as riskier or more volatile. 

From a toolkit perspective, we will accomplish the following in this section:

1. calculate and visualize x and rolling x where x is one of the following:
   - standard deviation 
   - skewness
   - kurtosis 
2. Build shiny apps for them 

# Standard Deviation

Here is the textbook equation for the standard deviation of a multi asset portfolio 

Standard Deviation = $\sqrt{\sum_{t = 1}^n (x_i - \bar{x})^2 / n}$

where x is each monthly portfolio rerturn, x bar is the mean monthly portfolio return and n is the number of observations. 

```{r}
# build a covariance matrix of returns
(cov_mat <- cov(asset_returns_xts) %>% round(5))

# take sqrt of the transpose of the weights vector times the cov matrix times the weights vector
sqrt(t(weights) %*% cov_mat %*% weights) * 100  %>%
    round(2) -> sd_matrix_algebra
```

We will also be looking at rolling standard deviation. We can look at the average volatility for the entire life of the portfolio, but we also want to understand how that volatility has changed over time and behaved in different market conditions. 

## Standard Deviation in the xts World 

```{r}
asset_returns_xts %>% StdDev(weights = weights) * 100 %>% round(2)
```

## Rolling Standard Deviations in the xts World

```{r}
port_rolling_sd_xts <- rollapply(portfolio_returns_xts_rebalanced_monthly,
                                 FUN = sd,
                                 # assign length of rolling window
                                 width = 24) %>%
    na.omit()
```

## Standard Deviation in the Tidyverse 

```{r}
portfolio_returns_dplyr %>%
    summarise(
        sd = sd(returns) %>% round(4)) -> portfolio_sd_tidy
```

## Rolling Standard Deviations in the Tidyverse

```{r}
port_rolling_sd_tidy <- portfolio_returns_dplyr %>%
    mutate(rolling_sd = rollapply(returns,
                                  FUN = sd,
                                  width = 24,
                                  fill = NA)) %>%
    select(date, rolling_sd) %>%
    na.omit()
```

<!-- ## Rolling Standard Deviation with Tibbletime and the Tidyverse -->

<!-- tibbletime and its rollify function are built for time series analysis. -->

<!-- ```{r} -->
<!-- sd_roll_24 <- rollify(sd, window = 24) -->

<!-- portfolio_tq %>% -->
<!--     as_tbl_time(index = date) %>% -->
<!--     mutate(sd = sd_roll_24(returns)) -->
<!-- ``` -->

## Standard Deviation in the tidyquant World 

```{r}
portfolio_tq %>%
    tq_performance(Ra = returns,
                   Rb = NULL,
                   performance_fun = table.Stats) %>%
    select(Stdev) %>%
    mutate(tq_sd = round(Stdev, 4) * 100)
```

## Rolling Standard Deviaton in the tidyquant World 

```{r}
port_rolling_sd_tq <- portfolio_tq %>%
    tq_mutate(mutate_fun = rollapply,
              width = 24,
              FUN = sd,
              col_rename = "rolling_sd") %>%
    select(date, rolling_sd) %>%
    na.omit()

port_rolling_sd_tq %>% tail(3)
```

# Visualizing Standard Deviation 

Visualizing standard deviation of portfolio returns comes down to visualizing the dispersion of portfolio returns. 

```{r}
portfolio_returns_dplyr %>%
    ggplot(aes(x = date, y = returns)) +
    geom_point(color = "mediumpurple") +
    scale_x_date() +
    ggtitle("Scatterplot of Returns by Date")
```

It is hard to tell which years were good and bad. Lets add a different color for any monthly returns that are one standard deviation away from the mean. 

```{r}
# get statistics
sd_plot <- sd(portfolio_tq$returns)
mean_plot <- mean(portfolio_tq$returns)
sd_up <- mean_plot + sd_plot
sd_down <- mean_plot - sd_plot

portfolio_tq %>%
    mutate(hist_red = if_else(returns < sd_down,
                              returns,
                              as.numeric(NA)),
           hist_green = if_else(returns > sd_down,
                                returns,
                                as.numeric(NA)),
           hist_blue = if_else(returns > sd_down &
                               returns < sd_up,
                               returns,
                               as.numeric(NA))) %>%
    ggplot(aes(x = date)) +
    geom_point(aes(y = hist_red), color = "red") +
    geom_point(aes(y = hist_green), color = "forestgreen") +
    geom_point(aes(y = hist_blue), color = "blue") +
    geom_hline(yintercept = sd_up,
               color = "mediumpurple",
               lty = 4) +
    geom_hline(yintercept = sd_down,
               color = "mediumpurple",
               lty = 4) + 
    labs(title = "Colored Scatter", y = "monthly returns") +
    scale_x_date()
```

To visualize the actual standard deviation of our portfolio, it helps to do so in a comparative manner. We can explore how our portfolio risk compares to the risk of the 5 individual assets. 

```{r}
asset_returns_long %>%
    group_by(asset) %>%
    summarise(sd = 100 * sd(returns)) %>%
    add_row(asset = "Portfolio",
            sd = portfolio_sd_tidy[[1]] * 100) %>%
    ggplot(aes(x = asset,
               y = sd,
               color = asset)) +
    geom_point() +
    scale_y_continuous(labels = function(x) paste0(x, "%")) +
    geom_text(aes(x = "Portfolio",
                  y = portfolio_sd_tidy[[1]] * 100 + 0.2),
              label = "Portfolio",
              color = "mediumpurple") +
    labs(y = "Standard Deviation")
```

We can also visualize expected monthly returns scattered against the standard deviation of monthly returns. 

```{r}
asset_returns_long %>%
    group_by(asset) %>%
    summarise(expected_return = mean(returns),
              std_dev = sd(returns)) %>%
    add_row(asset = "Portfolio",
            std_dev = sd(portfolio_tq$returns),
            expected_return = mean(portfolio_tq$returns)) %>%
    ggplot(aes(x = std_dev,
               y = expected_return,
               color = asset)) +
    geom_point(size = 2) +
    geom_text(aes(x = sd(portfolio_tq$returns) * 1.1,
                  y = mean(portfolio_tq$returns),
                  label = "Portfolio")) +
    ylab("Expected Return") +
    xlab("Standard Deviation") +
    ggtitle("Expected Monthly Return vs. Risk") +
    scale_y_continuous(labels = function(x) paste0(x, "%"))
```

# Visualizing Rolling Standard Deviation in the xts World 

```{r}
port_rolling_sd_xts %<>%
    round(4) * 100
```

```{r}
highchart(type = "stock") %>%
    hc_title(text = "24 Month Rolling Volatility") %>%
    hc_add_series(port_rolling_sd_xts,
                  color = "mediumpurple") %>%
    hc_add_theme(hc_theme_flat()) %>%
    hc_yAxis(labels = list(format = "{value}%"),
             opposite = FALSE) %>%
    hc_navigator(enabled = FALSE) %>%
    hc_scrollbar(enabled = FALSE) %>%
    hc_exporting(enabled = TRUE) %>%
    hc_legend(enabled = TRUE)
```

## Visualizing Rolling Standard Deviation in the Tidyverse 

```{r}
port_rolling_sd_tq %>%
    ggplot(aes(x = date)) +
    geom_line(aes(y = rolling_sd),
              color = "mediumpurple") +
    scale_y_continuous(labels = scales::percent) +
    scale_x_date() +
    labs(title = "Rolling Standard Deviation", y = "")
```

# Shiny App Standard Deviation 

Now we can wrap all that work intp an app that allows the user to choose a 5 asset portfolio and chart rolling volatility of different widths 

```{r}
symbols

#sym_in <- symbols %>% map2(., 1:5, ~ make_input(.x, .y))

date_input

window_in <- fluidRow(column(5,
                             numericInput("window",
                                          "window",
                                          value = 12,
                                          min = 3,
                                          max = 36,
                                          step = 1)))

ui <- fluidPage(
    theme = shinythemes::shinytheme("flatly"),
    sidebarLayout(
        sidebarPanel(!!!sym_in,
                     date_input,
                     window_in,
                     actionButton("go", "Submit")),
        mainPanel(## fluidRow(plotOutput("plot_1")),
                  fluidRow(plotOutput("plot_2")))))
```

```{r}
server <- function(input, output, session) {
    prices <- eventReactive(input$go, {
       syms <- c(input$stock_SPY, input$stock_EFA,
                 input$stock_IJS, input$stock_EEM, 
                 input$stock_AGG)

        prices <- getSymbols(syms,
                             src = "yahoo",
                             from = input$date,
                             auto.assign = TRUE,
                             warnings = FALSE) %>%
            map(., ~ Ad(get(.x))) %>%
            reduce(merge) %>%
            `colnames<-`(syms) })

    port_rolling_sd <- eventReactive(input$go, {
        weights <- c(input$wt_1 / 100, input$wt_2 / 100,
                    input$wt_3 / 100, input$wt_4 / 100,
                    input$wt_5 / 100)

        port_rolling <- prices() %>%
            to.monthly(indexAt = "last",
                       OHLC = FALSE) %>%
            tk_tbl(preserve_index = TRUE,
                   rename_index = "date") %>%
            slice(-1) %>%
            gather(asset, returns, -date) %>%
            group_by(asset) %>%
            mutate(returns = (log(returns) - log(lag(returns)))) %>% 
            tq_portfolio(assets_col = asset,
                         returns_col = returns,
                         weights = weights,
                         col_rename = "returns",
                         rebalance_on = "months")

        window <- input$window
        
        port_rolling_sd <- port_rolling %>%
            tq_mutate(mutate_fun = rollapply,
                      width = window,
                      FUN = sd,
                      col_rename = "rolling_sd") %>%
            select(date, rolling_sd) %>%
            na.omit()
    })

    output$plot_1 <- renderHighchart({
        port_rolling_sd() %>%
            tk_xts(date_col = date) %>%
            round(4) * 100 -> hc_tbl

        highchart(type = "stock") %>%
            hc_title(text = "Portfolio Rolling Volatility") %>%
            hc_yAxis(title = list("Volatility"),
                     labels = list(format = "{value}%"),
                     opposite = FALSE) %>%
            hc_add_series(hc_tbl,
                          name = "Portfolio Volatility",
                          color = "mediumpurple") %>%
            hc_add_theme(hc_theme_flat()) %>%
            hc_navigator(enabled = FALSE) %>%
            hc_scrollbar(enabled = FALSE) %>%
            hc_exporting(enabled = TRUE)
    })

    output$plot_2 <- renderPlot({
        port_rolling_sd() %>%
            ggplot(aes(x = date)) +
            geom_line(aes(y = rolling_sd),
                      color = "mediumpurple") +
            scale_y_continuous(labels = scales::percent) +
            ggtitle("Portfolio Rolling Volatility") +
            scale_x_date()})
}

shinyApp(ui, server)
```

# Skewness

Skewness is the degree to which returns are asymmetric around their mean. Skewness can be taken as a measure of how returns are not distributed normally.

Skew = $\frac{\sum_{t = 1}^n (x_i - \bar{x})^3}{(\sum_{t = 1}^n (x_i - \bar{x})^2 / n)^{3/2}}$

## Skewness in the xts World

```{r}
(skew_xts <- skewness(portfolio_returns_xts_rebalanced_monthly$returns))
```

## Skewness in the Tidyverse

```{r}
(skew_tidy <- portfolio_tq %>%
    summarize(skew_builtin = skewness(returns)))
```

## Visualizing Skewness

```{r}
portfolio_tq %>%
    ggplot(aes(x = returns)) +
    geom_histogram(alpha = 0.7,
                   binwidth = .003,
                   fill = "mediumpurple",
                   color = "black") +
    geom_vline(xintercept = 0, lty = 2)
```


 We can add color for the returns which are more than 2 standard deviations below the mean
 
```{r}
portfolio_tq %>%
    mutate(hist_red = if_else(returns < (mean(returns) - 2 * sd(returns)),
                              returns,
                              as.numeric(NA)),
           reg_returns = if_else(returns > (mean(returns) - 2 * sd(returns)),
                                 returns,
                                 as.numeric(NA))) %>%
    ggplot() +
    geom_histogram(aes(x = hist_red),
                   alpha = 0.7,
                   binwidth = 0.003,
                   fill = "red",
                   color = "black") +
    geom_histogram(aes(x = reg_returns),
                   alpha = 0.7,
                   binwidth = 0.003,
                   fill = "mediumpurple",
                   color = "black") +
    geom_vline(xintercept = 0, lty = 2) +
    xlab("Monthly Returns")
```

### Density 

```{r}
density_mm_plot <- function(data, x_val,
                     mean = TRUE,
                     median = TRUE,
                     sd = TRUE,
                     shade_fill = "gold",
                     shade_fill_med = "red") {
    # baseline
    data %>%
        ggplot(aes(x = .data[[x_val]])) +
        stat_density(geom = "line",
                     alpha = 1,
                     color = "black") -> p_dens_plot

    if (mean) {
        mean <- mean(data[[x_val]])
        
        # get area less than mean
        shaded_area_data <- ggplot_build(p_dens_plot)$data[[1]] %>%
                                                    filter(x < mean)

        p_dens_plot +
            geom_area(data = shaded_area_data,
                      aes(x = x, y = y),
                      fill = shade_fill,
                      alpha = 0.3) -> p_dens_plot

        p_dens_plot +
            geom_segment(data = shaded_area_data,
                         aes(x = mean,
                             y = 0,
                             xend = mean,
                             yend = density),
                         color = "black",
                         lty = 2,
                         alpha = 0.5) +
            annotate(geom = "text",
                     x = mean,
                     y = 5,
                     label = "mean",
                     color = "black",
                     fontface = "plain",
                     angle = 90,
                     alpha = 0.8,
                     vjust = -1.75) -> p_dens_plot
        }
    
    if (median) {
        median <- median(data[[x_val]])
        median_line_data <- ggplot_build(p_dens_plot)$data[[1]] %>%
                                                    filter(x <= median)

        p_dens_plot +
            geom_area(data = median_line_data,
                      aes(x = x, y = y),
                      fill = shade_fill_med,
                      alpha = 0.3) -> p_dens_plot

        p_dens_plot +
            geom_segment(data = median_line_data,
                         aes(x = median,
                             y = 0,
                             xend = median,
                             yend = density),
                         color = "black",
                         lty = 2,
                         alpha = 0.5) +
            annotate(geom = "text",
                     x = median,
                     y = 5,
                     label = "median",
                     color = "black",
                     fontface = "plain",
                     angle = 90,
                     alpha = 0.8,
                     vjust = 1.75) -> p_dens_plot
    }

    if (sd) {
        mean <- mean(data[[x_val]])

        sd_out <- c(mean - (2 * sd(data[[x_val]])),
                    mean + (2 * sd(data[[x_val]])))

        sd_shaded <- ggplot_build(p_dens_plot)$data[[1]] %>%
                                             split(between(.$x, sd_out[[1]],
                                                            sd_out[[2]])) %>%
                                             .[[1]] %>%
                                             split(.$x < sd_out[[1]])

        p_dens_plot +
            geom_segment(data = sd_shaded[[1]],
                         aes(x = sd_out[[2]],
                             y = 0,
                             xend = sd_out[[2]],
                             yend = density)) +
            geom_segment(data = sd_shaded[[2]],
                         aes(x = sd_out[[1]],
                             y = 0,
                             xend = sd_out[[1]],
                             yend = density)) +
            geom_area(data = sd_shaded[[2]],
                      aes(x = x, y = y),
                      fill = "blue",
                      alpha = 0.5) +
            geom_area(data = sd_shaded[[1]],
                      aes(x = x, y = y),
                      fill = "blue",
                      alpha = 0.5) -> p_dens_plot
    }

    p_dens_plot +
        geom_vline(xintercept = 0, lty = 2, alpha = 0.3) +
        xlab(x_val)
}

density_mm_plot(portfolio_tq, "returns")
```

Now we can investigate our portfolio's skewness compared to the 5 assets' skewness. 

```{r}
asset_returns_long %>%
    summarize(skew_assets = skewness(returns)) %>%
    add_row(asset = "Portfolio",
            skew_assets = skew_tidy$skew_builtin) %>%
    ggplot(aes(x = asset,
               y = skew_assets,
               color = asset)) +
    geom_point() +
    annotate(geom = "text",
             x = "Portfolio",
             y = skew_tidy$skew_builtin + 0.04,
             label = "Portfolio",
             color = "cornflowerblue") +
    labs(y = "skewness")
```

# Rolling Skewness in the xts World

```{r}
window <- 24
rollapply(portfolio_returns_xts_rebalanced_monthly,
          FUN = skewness,
          width = window) %>%
    na.omit() -> rolling_skew_xts
```

# Rolling Skewness in the Tidyverse with tibbletime

```{r}
# create a rolling function
skew_roll_24 <- rollify(skewness, window = window)

portfolio_tq %>%
    as_tbl_time(index = date) %>%
    mutate(skew = skew_roll_24(returns)) %>%
    select(-returns) %>%
    na.omit() -> rolling_skew_tbltime
```

# Rolling Skewness in the tidyquant World

```{r}
portfolio_tq %>%
    tq_mutate(select = returns,
              mutate_fun = rollapply,
              width = window,
              FUN = skewness,
              col_rename = "tq") %>%
    select(-returns) %>% 
    na.omit() -> rolling_skew_tq
```

Confirm that our results are consistent

```{r}
rolling_skew_tq %>%
    mutate(xts = coredata(rolling_skew_xts),
           tbltime = rolling_skew_tbltime$skew) %>%
    mutate_if(is.numeric, ~ round(.x, 3)) %>%
    tail(3)
```

# Visualizing Rolling Skewness

```{r}
highchart(type = "stock") %>%
    hc_title(text = "Rolling 24 Month Skewness") %>%
    hc_add_series(rolling_skew_xts,
                  name = "Rolling Skewness",
                  color = "mediumpurple") %>%
    hc_yAxis(title = list(text = "skewness"),
             opposite = FALSE,
             max = 1,
             min = -1) %>%
    hc_navigator(enabled = FALSE) %>%
    hc_scrollbar(enabled = FALSE) %>%
    hc_add_theme(hc_theme_flat()) %>%
    hc_exporting(enabled = TRUE)
```

```{r}
rolling_skew_tq %>%
    ggplot(aes(x = date, y = tq)) +
    geom_line(color = "mediumpurple") +
    ggtitle("Rolling 24 Month Skew") +
    ylab(paste("Rolling", window, "Month Skewness")) +
    scale_y_continuous(limits = c(-1, 1))
```

# Kurtosis

Kurtosis is a measure of the degree to which portfolio returns appear in the tails of their distribution. A normal distribution has a kurtosis of 3. A distribution with a kurtosis greater than 3 has more returns in its tails than the normal, and one with a kurtosis less than 3 has fewer returns in its tails than the normal.

Kurtosis is often described as negative excess or positive excess compared to a kurtosis of 3. A distribution with negative excess kurtosis equal to -1 has an absolute kurtosis of 2, because we subtract from 3. Negative kurtosis generally means less risk.

Kurtosis = $\frac{\sum_{t = 1}^n (x - \bar{x})^4 / n}{(\sum_{t = 1}^n (x_i - \bar{x})^2 - 3)}$

## Kurtosis in the xts world

```{r}
kurt_xts <- kurtosis(portfolio_returns_xts_rebalanced_monthly$returns)
```

## Kurtosis in the Tidyverse

```{r}
portfolio_tq %>%
    summarize(kurt = kurtosis(returns)) -> kurt_tidy
```

# Visualizing Kurtosis

```{r}
density_mm_plot(portfolio_tq, "returns")
```

We can also compare portfolio kurtosis to the individual assets kurtosis

```{r}
asset_returns_long %>%
    summarize(kurt_assets = kurtosis(returns)) %>%
    add_row(asset = "Portfolio",
            kurt_assets = kurt_tidy$kurt) %>%
    ggplot(aes(x = asset, y = kurt_assets, color = asset)) +
    geom_point() +
    annotate(geom = "text",
             x = "Portfolio",
             y = kurt_tidy$kurt + .06,
             label = "Portfolio",
             color = "cornflowerblue")
```

# Rolling Kurtosis in the xts World

```{r}
window <- 24

(rolling_kurt_xts <- rollapply(portfolio_returns_xts_rebalanced_monthly,
                              FUN = kurtosis,
                              width = window) %>%
    na.omit())
```

# Rolling Kurtosis in the Tidyverse with tibbletime

```{r}
kurt_roll_24 <- rollify(kurtosis, window = window)

(roll_kurt_tbl_time <- portfolio_tq %>%
    as_tbl_time(index = date) %>%
    mutate(kurt = kurt_roll_24(returns)) %>%
    select(-returns) %>%
    na.omit())
```

# Rolling Kurtosis in the tidyquant World

```{r}
(rolling_kurt_tq <- portfolio_tq %>%
    tq_mutate(select = returns,
              mutate_fun = rollapply,
              width = window,
              FUN = kurtosis,
              col_rename = "tq") %>%
    select(-returns) %>%
    na.omit())
```

# Visualizing Rolling Kurtosis

```{r}
highchart(type = "stock") %>%
    hc_title(text = "Rolling 24 Month Kurtosis") %>%
    hc_add_series(rolling_kurt_xts,
                  name = "Rolling 24 Month Kurtosis",
                  color = "mediumpurple") %>%
    hc_yAxis(title = list(text = "kurtosis"),
             opposite = FALSE) %>%
    hc_add_theme(hc_theme_flat()) %>%
    hc_navigator(enabled = FALSE) %>%
    hc_scrollbar(enabled = FALSE) %>%
    hc_exporting(enabled = TRUE)
```

```{r}
rolling_kurt_tq %>%
    ggplot(aes(x = date, y = tq)) +
    geom_line(color = "mediumpurple") +
    ggtitle("Rolling 24 Month Kurtosis") +
    labs(y = "Rolling Kurtosis")
```

# Shiny App Skewness and Kurtosis

```{r}

symbols <- c("SPY", "EFA", "IJS", "EEM", "AGG")

make_input <- function(symbol, index) {
    fluidRow(
        column(6,
               textInput(paste0("stock_", symbol),
                         paste("Stock", index),
                         symbol)),
        column(6,
               numericInput(paste0("wt_", index),
                            "Portfolio %",
                            value = 20,
                            min = 0,
                            max = 100)))
}

stock_inputs <- imap(symbols, ~ make_input(.x, .y))

date_input <- fluidRow(
    column(6,
           dateInput("date",
                     "Starting Date",
                     "2013-01-01",
                     format = "yyyy-mm-dd")),
    column(6,
           numericInput("window",
                        "Window",
                        min = 1,
                        max = 24,
                        value = 6)))

buttons <- fluidRow(
    column(6,
           actionButton("roll",
                        "Rolling",
                        class = "btn btn-primary")),
    column(6,
           actionButton("returns",
                        "Returns",
                        class = "btn btn-success")))

ui <- fluidPage(
    theme = shinythemes::shinytheme("paper"),
    sidebarLayout(
        sidebarPanel(
            !!!stock_inputs,
            date_input,
            br(),
            buttons),
        mainPanel(
            tabsetPanel(
                tabPanel("Rolling Skewness",
                         highchartOutput("plot_skew")),
                tabPanel("Rolling Kurtosis",
                         highchartOutput("plot_kurt"))),
            tabsetPanel(
                tabPanel("Number Threshold"),
                tabPanel("Std Dev Threshold"),
                tabPanel("Density Mean"),
                tabPanel("Density Tails")))))
```

```{r}
get_prices <- function(syms, from_date) {
    getSymbols(syms,
               src = "yahoo",
               from = from_date,
               auto.assign = TRUE,
               warnings = FALSE) %>%
        map(., ~ Ad(get(.x))) %>%
        reduce(merge) %>%
        `colnames<-`(syms)    
}

roll_return <- function(data, fun, window) {
        rollapply(data,
                  FUN = fun,
                  width = window) %>%
            na.omit()
}

hchart_plot <- function(data, title) {
    highchart(type = "stock") %>%
        hc_title(text = title) %>%
        hc_add_series(data,
                      name = title,
                      color = "mediumpurple") %>%
        hc_yAxis(title = str_extract(title, "[A-Za-z]*$"),
                 opposite = FALSE,
                 max = 3,
                 min = -3) %>%
        hc_navigator(enabled = FALSE) %>%
        hc_scrollbar(enabled = FALSE) %>%
        hc_exporting(enabled = TRUE)
}

server <- function(input, output, session) {
    ## Rolling Skewness and Kurtosis
    ########################################################
    # calculate asset returns
    prices <- eventReactive(input$roll, {
        # place parameters in vectors
        syms <- c(input$stock_SPY, input$stock_EFA,
                  input$stock_IJS, input$stock_EEM,
                  input$stock_AGG)

        weights <- c(input$wt_1 / 100, input$wt_2 / 100,
                     input$wt_3 / 100, input$wt_4 / 100,
                     input$wt_5 / 100)
        
        # get prices of the symbols
        get_prices(syms, input$date)})
    
    # calculate portfolio returns
    xts_returns <- eventReactive(input$roll, {
        prices() %>%
            to.monthly(indexAt = "last", OHLC = FALSE) %>%
            na.omit(Return.calculate(method = "log")) %>%
            Return.portfolio(weights = weights) %>%
            `colnames<-`("returns")})
    
    # calculate rolling skewness
    rolling_skew <- eventReactive(input$roll, {
        xts_returns() %>%
            roll_return(skewness, input$window)})
    
    # calculate rolling kurtosis
    rolling_kurtosis <- eventReactive(input$roll, {
        xts_returns() %>%
            roll_return(kurtosis, input$window)})
    
    # render plots
    output$plot_skew <- renderHighchart({
        hchart_plot(rolling_skew(),
                    "Rolling Skew")})

    output$plot_kurt <- renderHighchart({
        hchart_plot(rolling_kurtosis(),
                    "Rolling Kurtosis")})
}

shinyApp(ui, server)
```
